---
title: "Case Study 1"
author: "OLP"
date: "2023-03-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

cantidad de member y casual

## Ask
### Task business

#### Design marketing strategies aimed at converting casual riders into annual members. 

In order to do that, however, the marketing analyst team needs to better understand how annual members and casual riders differ, why casual riders would buy a membership, and how digital media could affect their marketing tactics. 

Three questions will guide the future marketing program:
1. How do annual members and casual riders use Cyclistic bikes differently?
2. Why would casual riders buy Cyclistic annual memberships?
3. How can Cyclistic use digital media to influence casual riders to become members?

The problem is how do annual members and casual riders use Cyclistic bikes differently.

The insights will indicate the trends of each member

## Prepare
La fuente de datos utilizada para el proyecto es de <https://divvy-tripdata.s3.amazonaws.com/index.html>  


The data has been made available by Motivate International Inc. under this license <https://ride.divvybikes.com/data-license-agreement>  


The data is zip and is organized in CSV files in comma-separated long format.
ROCCC - Reliable Original Comprehensive Current Cited  


The downloaded data is from February 2022 to January 2023.  

1 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
2 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
3 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
4 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
6 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
7 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
8 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
9 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
10 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
11 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
12 file : 202301-divvy-tripdata.zip - 202301-divvy-tripdata.csv  

Los pasos fueron los siguientes:  
1- Ubicar los 12 archivos en una carpeta  
2- Leer los 12 archivos en un solo dataframe, para ello deben tener la misma estructura  
3- Guardar el dataframe es un solo archivo csv (Feb2022-Ene2023.csv) de 1.2GB aprox.
4- Importar el archivo Feb2022-Ene2023.csv a MS SQL debido al tamaño  
5- Se crea una tabla Tripdata_2022 equivalente al Feb2022-Ene2023.csv  
6- Generar querys y exportarlas a formato csv  

Se generaron 9 querys:  

freq_by_trip_for_member_for_month.csv  
duration_trip_casual.csv  
duration_trip_member.csv  
end_station_by_member_top100.csv  
start_station_by_member_top100.csv  
freq_by_trip_for_member_for_day.csv  
num_ride_eight_hours.csv  
cant_trip_per_month.csv  
cant_trip_per_day.csv

Para efectuar el analisis We have chosen R porque podemos documentar todo el proceso: datos, graficos y conclusiones.  

```{r load packages}
library(tidyverse)
library(skimr)
library(janitor)
library(lubridate)
```

Read dataset CSV
```{r read csv}
df_freq_trip_mem_month <- read.csv("freq_by_trip_for_member_for_month.csv",sep=';')
df_freq_trip_mem_day <- read.csv("freq_by_trip_for_member_for_day.csv",sep=';')
df_dur_trip_casual <- read.csv("duration_trip_casual.csv",sep=';') 
df_dur_trip_member <- read.csv("duration_trip_member.csv",sep=';') 
df_end_stat_top100 <- read.csv("end_station_by_member_top100.csv",sep=';')  
df_start_stat_top100 <- read.csv("start_station_by_member_top100.csv",sep=';') 
df_num_ride_hours <- read.csv("num_ride_eight_hours.csv",sep=';')  
df_num_trip_month <- read.csv("cant_trip_per_month.csv",sep=';')
df_num_trip_day <- read.csv("cant_trip_per_day.csv",sep=';')
```


We explore dataframes
```{r Number of trip for member por month}
glimpse(df_freq_trip_mem_month)
```
```{r Number of trip for member for day of week}
glimpse(df_freq_trip_mem_day)
```
```{r trip duration of casual members greater than eight hours}
glimpse(df_dur_trip_casual)
```
```{r trip duration of annual members greater than eight hours}
glimpse(df_dur_trip_member)
```
```{r the top 100 most used start stations}
glimpse(df_start_stat_top100)
```
```{r the top 100 most used end stations}
glimpse(df_end_stat_top100)
```

```{r number of trips of more than 8 hours per member}
glimpse(df_num_ride_hours)
```

```{r number of trips by month per member}
glimpse(df_num_trip_month)
```

```{r number of trips by day of week per member}
glimpse(df_num_trip_day)
```
