---
title: "Case Study 1"
author: "OLP"
date: "2023-03-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

cantidad de member y casual

## Ask
### Task business

#### Design marketing strategies aimed at converting casual riders into annual members. 

In order to do that, however, the marketing analyst team needs to better understand how annual members and casual riders differ, why casual riders would buy a membership, and how digital media could affect their marketing tactics. 

Three questions will guide the future marketing program:
1. How do annual members and casual riders use Cyclistic bikes differently?
2. Why would casual riders buy Cyclistic annual memberships?
3. How can Cyclistic use digital media to influence casual riders to become members?

The problem is how do annual members and casual riders use Cyclistic bikes differently.

The insights will indicate the trends of each member

## Prepare
La fuente de datos utilizada para el proyecto es de <https://divvy-tripdata.s3.amazonaws.com/index.html>  


The data has been made available by Motivate International Inc. under this license <https://ride.divvybikes.com/data-license-agreement>  


The data is zip and is organized in CSV files in comma-separated long format.
ROCCC - Reliable Original Comprehensive Current Cited  


The downloaded data is from February 2022 to January 2023.  

1 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
2 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
3 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
4 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
6 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
7 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
8 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
9 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
10 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
11 file : 202202-divvy-tripdata.zip - 202202-divvy-tripdata.csv  
12 file : 202301-divvy-tripdata.zip - 202301-divvy-tripdata.csv  

Los pasos fueron los siguientes:  
1- Ubicar los 12 archivos en una carpeta  
2- Leer los 12 archivos en un solo dataframe, para ello deben tener la misma estructura  
3- Guardar el dataframe es un solo archivo csv (Feb2022-Ene2023.csv) de 1.2GB aprox.
4- Importar el archivo Feb2022-Ene2023.csv a MS SQL debido al tamaño  
5- Se crea una tabla Tripdata_2022 equivalente al Feb2022-Ene2023.csv  
6- Generar querys y exportarlas a formato csv  

Se generaron 9 querys:  

freq_by_trip_for_member_for_month.csv  
duration_trip_casual.csv  
duration_trip_member.csv  
end_station_by_member_top100.csv  
start_station_by_member_top100.csv  
freq_by_trip_for_member_for_day.csv  
num_ride_eight_hours.csv  
cant_trip_per_month.csv  
cant_trip_per_day.csv

Para efectuar el analisis We have chosen R porque podemos documentar todo el proceso: datos, graficos y conclusiones.  

```{r load packages}
library(tidyverse)
library(skimr)
library(janitor)
library(lubridate)
```

Read dataset CSV
```{r read csv}
df_freq_trip_mem_month <- read.csv("freq_by_trip_for_member_for_month.csv",sep=';')
df_freq_trip_mem_day <- read.csv("freq_by_trip_for_member_for_day.csv",sep=';')
df_dur_trip_casual <- read.csv("duration_trip_casual.csv",sep=';') 
df_dur_trip_member <- read.csv("duration_trip_member.csv",sep=';') 
df_end_stat_top100 <- read.csv("end_station_by_member_top100.csv",sep=';')
df_start_stat_top100 <- read.csv("start_station_by_member_top100.csv",sep=';') 
df_num_ride_hours <- read.csv("num_ride_eight_hours.csv",sep=';')  
df_num_trip_month <- read.csv("cant_trip_per_month.csv",sep=';')
df_num_trip_day <- read.csv("cant_trip_per_day.csv",sep=';')
```


We explore dataframes
```{r Number of trip for member por month}
glimpse(df_freq_trip_mem_month)
```
```{r Number of trip for member for day of week}
glimpse(df_freq_trip_mem_day)
```
```{r trip duration of casual members greater than eight hours}
glimpse(df_dur_trip_casual)
```
```{r trip duration of annual members greater than eight hours}
glimpse(df_dur_trip_member)
```
```{r the top 100 most used start stations}
glimpse(df_start_stat_top100)
```
```{r the top 100 most used end stations}
glimpse(df_end_stat_top100)
```

```{r number of trips of more than 8 hours per member}
glimpse(df_num_ride_hours)
```

```{r number of trips by month per member}
glimpse(df_num_trip_month)
```

```{r number of trips by day of week per member}
glimpse(df_num_trip_day)
```
## Process
The decision is to join the 12 files into one, import it into a database and create sql queries to extract summarized information, then analyze that information with R to be able to document the entire process.  

To ensure data integrity, the download has been checked against the zip files in the web.

Find null and empty values.  

Querys:  
Select 
      count(*) 
      from Tripdata_2022 
      where start_station_id=''  
      
843525 empty values in start_station_id  


Select  
      count(*)  
      from Tripdata_2022  
      where end_station_id=''  
      
902655 empty values in end_station_id  


Select  
      count(*)  
      from Tripdata_2022  
      where member_casual=''

No empty values in member_casual  

There are only two values for the column member_casual: 'member' and 'casual'  

## Analize
We find insights and trends  

```{r arrange freq number trip member} 
arrange(df_freq_trip_mem_month, desc(num_trip))
```

### Plot of num trip vs month
```{r}
ggplot(data=df_freq_trip_mem_month, aes(x=Month, y=num_trip)) + 
  geom_point() + 
  geom_smooth() + 
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = seq(from = 1, to = 12, by = 1)) +
  ylab("Number trips") +
  xlab("Month") +
  facet_wrap(~member_casual) +
  labs(title="Number trips vs. Month")
```
Los meses que mas trips hay son de mayo a octubre para ambos miembros.  



```{r}
ggplot(data=df_freq_trip_mem_month, aes(x=Month, y=num_trip, 
                                        group=member_casual)) + 
  geom_line(aes (linetype=member_casual, color=member_casual), size=1.5) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = seq(from = 1, to = 12, by = 1)) +
  ylab("Number trips") +
  xlab("Month") +
  labs(title="Number trips vs. Month")
```



```{r}
ggplot(data=df_freq_trip_mem_day, aes(x=day_of_week, y=num_trip, 
                                        group=member_casual,
                                        color=member_casual)) + 
  geom_line(size=2) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = seq(from = 1, to = 7, by = 1)) +
  ylab("Number trips") +
  xlab("Day of week") +
  facet_wrap(~Month) +
  labs(title="Number trips by day of week")
```

Calcular el promedio, minimo y maximo de cantidad de trips para cadia dia de la semana.  


```{r average min max number trips} 
df_summ <- df_freq_trip_mem_day %>%
  group_by(member_casual, day_of_week) %>%
  summarise(average_num_trip=mean(num_trip),
            min_num_trip=min(num_trip),
            max_num_trip=max(num_trip))
#  filter(member_casual=="member") %>%
arrange(df_summ, desc(average_num_trip))
```


```{r plot average vs day of week} 
ggplot(data=df_summ, aes(x=day_of_week, y=average_num_trip, 
                                        group=member_casual,
                                        color=member_casual)) + 
  geom_line(size=2) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = seq(from = 1, to = 7, by = 1)) +
  ylab("Average trips") +
  xlab("Day of week 1:Sunday .. 7:Saturday") +
  labs(title="Average of trips vs day of week")
```

En el plot se observa que:  
Los miembros anuales viajan mas en promedio los dias martes a jueves.  
Los miembros casuales viajan mas en promedio os dias sabados y domingos
Esto es correcto si tenemos en cuenta que los miembros anuales usan Cyclistic para ir al trabajo y los miembros casual para pasear.


```{r}
df_num_trip_month %>%
  group_by(member_casual) %>%
  select(num_trip, member_casual) %>%
  summarise(total_trip_member = sum(num_trip))
```

Los miembros annual viajan 1.5 veces mas que los miembros casual.  


Find cantidad de trips para start station por member

```{r top10 start station number trips}
df_start_stat_top100 %>%
  arrange(desc(num_start_Station)) %>%
  slice_head(n = 10)
```

Las estaciones de salida mas usadas son:  
Streeter Dr & Grand Ave  
DuSable Lake Shore Dr & Monroe St  
Millennium Park  
Michigan Ave & Oak St  
Y son usadas por miembros casual


```{r top10 number trips end station}
df_end_stat_top100 %>%
  arrange(desc(num_end_Station)) %>%
  slice_head(n = 10)
```

Las estaciones de llegada mas usadas son:  
Streeter Dr & Grand Ave  
DuSable Lake Shore Dr & Monroe St  
Millennium Park  
Michigan Ave & Oak St  
DuSable Lake Shore Dr & North Blvd  
y son usadas por miembros casual

Resumen miembros casual  
La estación de salida y llegada mas usada para miembros casual es
Streeter Dr & Grand Ave con 58270 salidas y 60092 llegadas.


Resumen miembros annual  
La estacion de salida y llegada mas usada es Kingsbury St & Kinzie St con 25239
trips para la salida y 25094 para la llegada.  


Find trips superior a ocho horas

```{r}
view(df_num_ride_hours)
```


Los miembros casual tienen cuatro veces viajes mas largos que los
miembros annual.
